\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-2\baselineskip}

### 2017-05-31 paper/ms.tex
Added text in simulation section.

K &=& \\
&&\begin{pmatrix}

@@

>>>>

K &=&

@@

Simulation of a Gaussian process for large $N$ requires multiplying an
$N$-dimensional vector of Gaussian normal deviates, $\bf{g}$, by the square root
of the covariance matrix:
\begin{equation}
{\bf y} = K_\bvec{\alpha}^{1/2} {\bf g},
\end{equation}
where $g_i \sim N(0,1)$.  Note that $K_\bvec{\alpha} = K_\bvec{\alpha}^{1/2} \left(K_\bvec{\alpha}^{1/2}\right)^T$, which
comparing with the Cholesky factorization gives $K_\bvec{\alpha}^{1/2} = L D^{1/2}$,
where $(D^{1/2})_{ii} = \sqrt{D_{ii}}$.  Also note that the exponent of
the likelihood with $\mu_\theta =0$ is given by
\begin{eqnarray}
Q &=&-\frac{1}{2}{\bf y}^T K_\bvec{\alpha}^{-1} {\bf y}\\
&=& -\frac{1}{2}{\bf y}^T \left(K_\bvec{\alpha}^{-1/2}\right)^T K_\bvec{\alpha}^{-1/2} {\bf y}\\
&=& -\frac{1}{2}{\bf g}^T K_\bvec{\alpha}^{1/2} \left(K_\bvec{\alpha}^{-1/2}\right)^T K_\bvec{\alpha}^{-1/2} K_\bvec{\alpha}^{1/2} {\bf g}\\
&=& -\frac{1}{2}{\bf g}^T {\bf g}.
\end{eqnarray}
The last quantity is simply the dot product of an $N$-dimensional vector of independent standard normal random variables with itself,
which is distributed according to the chi-squared distribution with $N$ degrees of freedom, i.e.\ $Q \sim \chi_N^2$.

To generate a draw from a Gaussian process with a specified \celerite\ model requires carrying out the
low-rank Cholesky decomposition first, generating the vector ${\bf g}$ of random normal deviates, and then multiplying
\begin{equation}
{\bf y} = L D^{1/2} {\bf g}.
\end{equation}
This may be carried out in $O(JN)$ operations with
\begin{eqnarray}
f_{n,j} &=& \phi_{n-1,j} \left(f_{n-1,j} + \tilde{X}_{n-1,j} g_{n-1}\right)\\
y_n &=& D_{n,n}^{1/2} g_n + \sum_{k=1}^{2J} \tilde{U}_{n,k} f_{n,k}
\end{eqnarray}
for $n = 1$ to $N$ and $j=1$ to $2J$, with $f_{0,j} = 0$ for all $j$ and $g_0 = 0$.
We expect that this simulation algorithm will be useful for testing Gaussian
process inference algorithms with large datasets, which becomes prohibitive
for large $N$ if standard Cholesky decomposition and multiplication are used.

\subsection{Prediction}
